Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-06 paper/main.tex
Updates
updates

\begin{figure}[t!]
    \begin{subfigure}{0.25 \textwidth}
    	\includegraphics[page=1]{figures.pdf}
    	\caption{$ N = 400 $}
    \end{subfigure}%
    \begin{subfigure}{0.25 \textwidth}
    	\includegraphics[page=2]{figures.pdf}
    	\caption{$ N = 800$ }
    \end{subfigure}
    \begin{subfigure}{0.25 \textwidth}
    	\includegraphics[page=5]{figures.pdf}
    	\caption{$ N = 400 $ }
    \end{subfigure}%
    \begin{subfigure}{0.25\textwidth}
    	\includegraphics[page=6]{figures.pdf}
    	\caption{ $ N = 800$ }
    \end{subfigure}
    \begin{subfigure}{0.25 \textwidth}
    	\includegraphics[page=3]{figures.pdf}
    	\caption{ $ N = 400$ }
    \end{subfigure}%
    \begin{subfigure}{0.25\textwidth}
    	\includegraphics[page=4]{figures.pdf}
    	\caption{$ N = 800$ }
    \end{subfigure}

@@

\section{Background and Related Work}
\label{sec:background}
\subsection{Notations} 
\todo[inline]{Need to rewrite this}
We denote random variables with uppercase letters $ X $, and a set of random
variables as $ \bm{X} = \{X_1, \cdots, X_k\} $ with $ \rvert \bm{X} \rvert = k
$. A sample from the random variable $ X $ is denoted as $ x $ and from a set
of random variables $ \bm{X} $ is denoted as $ \bm{x} $. In this paper, we
consider random variables in the mixed data setting where each of the random
variables can be either of continuous, categorical, or ordinal unless
specified. We write the expectation of a variable $ X $ as $ \mathbb{E}[X] $,
conditional expectations as $ \mathbb{E}[X | \bm{Z}] $. We denote a graph $ G =
(V, E) $ which nodes $ V$ and edges $ E $. In this paper we focus on all
observed variables, and linear measure of association. 
where $ \mathrm{cov}(X, Y) $ is the covariance between $ X $ and $ Y $, $ \mathrm{corr}(X, Y) $ is the correlation between $ X $ and $ Y $, and $
\sigma_X $ and $ \sigma_Y $ are the standard deviations of $ X $ and $ Y $
respectively.  Covariance matrix using $ \Sigma $ and the entry corresponding to
covariance between $ X $ and $ Y $ in the matrix using $ \Sigma_{XY} $.

@@

\begin{enumerate}
	\item \textbf{Continuous:} We train a model, $ E_X: x \sim
		\bm{z} $. The residuals are then computed by taking the difference
		between the true and the predicted values using $ E_X $. 
		$$ R_{x_i} = x_i - E_X(\bm{z}_i) $$
	\item \textbf{Ordinal:} We start by training a probability estimator, $
		p_X: x \sim \bm{z} $, and then use the estimated probabilities, 
		$ \hat{p}_X(x) $ to compute the residuals:
		$$ R_{x_i} = \hat{p}_X(X < x_i) - \hat{p}_X(X > x_i) $$
	\item \textbf{Categorical:} We again start by training a probability
		estimator $ p_X: x \sim \bm{z} $, and obtain probability
		estimates $ \hat{p}_X: p_X(\bm{z}) $. Next, we dummy encode the
		categorical variable and then compute the residuals as follows: 
		$$ R_{x_i} = x_i - \hat{p}(\bm{z}_i) $$
\end{enumerate}

@@

We repeat the above residualization step for both the variables $ X $ and $ Y $
obtaining residual matrices $ R_X $ and $ R_Y $. The type of variable determines
the shape of these matrices. If the variable is continuous or ordinal, the 
residual matrix is of shape $ (n \times 1 ) $ and if the variable is categorical,
we get a residual matrix of shape $ (x \times (k - 1)) $, where $ k $ is the number
of categories of the variable.

@@

\begin{equation}
	\rho(X, Y; \bm{Z}) = \frac{1}{\min(\rvert R_X \rvert, \rvert R_Y \rvert)}
	\sum_{i=1}^{\min(\rvert R_X \rvert, \rvert R_Y \rvert)} (\nu_i(R_X, R_Y))^2
\end{equation}

>>>>

\begin{figure}[t!]
    \begin{subfigure}{0.25 \textwidth}
	\centering
    	\includegraphics[page=1]{figures_v2.pdf}
    	\caption{$ N = 400 $}
    \end{subfigure}
    \begin{subfigure}{0.25\textwidth}
    	\includegraphics[page=2]{figures_v2.pdf}
    	\caption{ $ N = 800$ }
    \end{subfigure}
    \begin{subfigure}{0.25 \textwidth}
    	\includegraphics[page=3]{figures_v2.pdf}
    	\caption{ $ N = 400$ }
    \end{subfigure}%
    \begin{subfigure}{0.25\textwidth}
    	\includegraphics[page=4]{figures_v2.pdf}
    	\caption{$ N = 800$ }
    \end{subfigure}

@@

\section{Background and Related Work}
\label{sec:background}
\todo[inline]{Need to rewrite this}
We denote random variables with uppercase letters $ X $, and a set of random
variables as $ \bm{X} = \{X_1, \cdots, X_k\} $ with $ \rvert \bm{X} \rvert = k
$. A sample from the random variable $ X $ is denoted as $ x $ and from a set
of random variables $ \bm{X} $ is denoted as $ \bm{x} $. In this paper, we
consider random variables in the mixed data setting where each of the random
variables can be either of continuous, categorical, or ordinal unless
specified. We write the expectation of a variable $ X $ as $ \mathbb{E}[X] $,
conditional expectations as $ \mathbb{E}[X | \bm{Z}] $. We denote a graph $ G =
(V, E) $ which nodes $ V$ and edges $ E $. In this paper we focus on all
observed variables, and linear measure of association. 
where $ \mathrm{cov}(X, Y) $ is the covariance between $ X $ and $ Y $, $ \mathrm{corr}(X, Y) $ is the correlation between $ X $ and $ Y $, and $
\sigma_X $ and $ \sigma_Y $ are the standard deviations of $ X $ and $ Y $
respectively.  Covariance matrix using $ \Sigma $ and the entry corresponding to
covariance between $ X $ and $ Y $ in the matrix using $ \Sigma_{XY} $.

@@

\begin{enumerate}
	\item \textbf{Continuous:} We train a model, $ E_X: x \sim
		\bm{z} $. The residuals are then computed by taking the difference
		between the true and the predicted values using $ E_X $. 
		$$ R_{x_i} = x_i - E_X(\bm{z}_i) $$
	\item \textbf{Ordinal:} We start by training a probability estimator, $
		p_X: x \sim \bm{z} $, and then use the estimated probabilities, 
		$ \hat{p}_X(x) $ to compute the residuals:
		$$ R_{x_i} = \hat{p}_X(X < x_i) - \hat{p}_X(X > x_i) $$
	\item \textbf{Categorical:} We again start by training a probability
		estimator $ p_X: x \sim \bm{z} $, and obtain probability
		estimates $ \hat{p}_X: p_X(\bm{z}) $. Next, we dummy encode the
		categorical variable, resulting in a binary vector and then
		compute the residuals as follows: 
		$$ R_{x_i} = x_i - \hat{p}_X(\bm{z}_i) $$
\end{enumerate}

@@

We repeat the above residualization step for both the variables $ X $ and $ Y $
obtaining residual matrices $ R_x $ and $ R_y $. The type of variable determines
the shape of these matrices. If the variable is continuous or ordinal, the 
residual matrix is of shape $ (n \times 1 ) $ and if the variable is categorical,
we get a residual matrix of shape $ (x \times (k - 1)) $, where $ k $ is the number
of categories of the variable.

@@

\begin{equation}
	\rho(X, Y; \bm{Z}) = \frac{1}{\min(\rvert R_x \rvert, \rvert R_y \rvert)}
	\sum_{i=1}^{\min(\rvert R_x \rvert, \rvert R_y \rvert)} (\nu_i(R_x, R_y))^2
\end{equation}
